Rectangulotopes

### 2024-04-08 draft/rectangulotopes.tex
first facets

\begin{theorem}
\label{thm:weakRectangulotope}
  The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
  \begin{enumerate}[(i)]
  \item the convex hull of the points
  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i$ for all weak rectangulations $R$ of size~$n$,
  with
  \[
%  \begin{split}
%    \loday{w}^R_i \eqdef & h(T, i)\cdot v(T,i) \\
%    \antiloday{w}^R_i \eqdef & h(S, i)\cdot v(S,i),
%  \end{split}
    \loday{w}^R_i \eqdef h^T_i\cdot v^T_i
    \qquad\text{and}\qquad
    \antiloday{w}^R_i \eqdef h^S_i\cdot v^S_i,
  \]
  where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
  \end{enumerate}
\end{theorem}

For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
The coordinates of the vertices corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
\vincent{add table for the facets}

\begin{table}[p]
	\centerline{
	\begin{tabular}{cccccccc}
		\includegraphics[scale=1.5]{rectangulation1} &
		\includegraphics[scale=1.5]{rectangulation2} &
		\includegraphics[scale=1.5]{rectangulation3} &
		\includegraphics[scale=1.5]{rectangulation4} &
		\includegraphics[scale=1.5]{rectangulation5} &
		\includegraphics[scale=1.5]{rectangulation6} &
		\\
		$(-3, -1, 1, 3)$ &
		$(-3, -1, 5, -1)$ &
		$(-3, 3, -3, 3)$ &
		$(1, -5, 1, 3)$ &
		$(-3, 0, 5, -2)$ &
		$(-3, 5, -3, 1)$ &
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation8} &
		\includegraphics[scale=1.5]{rectangulation9} &
		\includegraphics[scale=1.5]{rectangulation10} &
		\includegraphics[scale=1.5]{rectangulation11} &
		\includegraphics[scale=1.5]{rectangulation12} &
		\includegraphics[scale=1.5]{rectangulation13} &
		\\
		$(-1, 3, -5, 3)$ &
		$(2, -5, 0, 3)$ &
		$(-3, 5, 0, -2)$ &
		$(-2, 0, 5, -3)$ &
		$(1, -5, 5, -1)$ &
		$(-1, 5, -5, 1)$ &
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation14} &
		\includegraphics[scale=1.5]{rectangulation15} &
		\includegraphics[scale=1.5]{rectangulation16} &
		\includegraphics[scale=1.5]{rectangulation17} &
		\includegraphics[scale=1.5]{rectangulation19} &
		\includegraphics[scale=1.5]{rectangulation20} &
		\\
		$(3, -5, 0, 2)$ &
		$(2, 0, -5, 3)$ &
		$(-2, 5, 0, -3)$ &
		$(1, -3, 5, -3)$ &
		$(3, -5, 3, -1)$ &
		$(3, 0, -5, 2)$ &
		\\[.2cm]
		&		
		\includegraphics[scale=1.5]{rectangulation21} &
		\includegraphics[scale=1.5]{rectangulation22} &
		\includegraphics[scale=1.5]{rectangulation23} &
		\includegraphics[scale=1.5]{rectangulation24} &
		\\
		&
		$(-1, 5, -1, -3)$ &
		$(3, -3, 3, -3)$ &
		$(3, 1, -5, 1)$ &
		$(3, 1, -1, -3)$ &
	\end{tabular}
	}
	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
	\label{tab:verticesWeakRectangulotope}
\end{table}

@@

\begin{theorem}
\label{thm:strongRectangulotope}
  The $(n-1)$-dimensional strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
  \begin{enumerate}[(i)]
  \item convex hull of the points
  \[
  \sum_{i,j\in [n], i< j} (\yin{w}^R_{i,j} - \yang{w}^R_{i,j})\cdot (\b{e}_i - \b{e}_j),
  \]
   for all strong rectangulations $R$ of size~$n$, with
  \[
%  \begin{split}
%    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
%    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket,
%  \end{split}
    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \, \backslash \, j \rrbracket
    \qquad\text{and}\qquad
    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \neg \, j \, \backslash \, i \rrbracket,
  \]
 where:
  \begin{itemize}
  \item $S$ and $T$ are the source and target trees of the rectangulation~$R$,
  \item $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
  \item $ch^{T,S}_{i,j}$ (resp.~$cv^{T,S}_{i,j}$) denote the number of common leaves of the horizontal (resp.~vertical) subtree of $i$ in $T$ and the horizontal (resp.~vertical) subtree of $j$ in $S$,
  \end{itemize}
  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = \binom{n+1}{2}$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
  \end{enumerate}
\end{theorem}

@@

For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
The coordinates of the vertices corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
\vincent{add table for the facets}

\begin{table}
	\centerline{
	\begin{tabular}{cccccccc}
		\includegraphics[scale=1.5]{rectangulation1} &
		\includegraphics[scale=1.5]{rectangulation2} &
		\includegraphics[scale=1.5]{rectangulation3} &
		\includegraphics[scale=1.5]{rectangulation4} &
		\includegraphics[scale=1.5]{rectangulation5} &
		\includegraphics[scale=1.5]{rectangulation6} &
		\\
		$(-6, -2, 2, 6)$ &
		$(-6, -2, 9, -1)$ &
		$(-6, 5, -5, 6)$ &
		$(1, -9, 2, 6)$ &
		$(-6, 1, 9, -4)$ &
		$(-6, 9, -5, 2)$ &
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation8} &
		\includegraphics[scale=1.5]{rectangulation9} &
		\includegraphics[scale=1.5]{rectangulation10} &
		\includegraphics[scale=1.5]{rectangulation11} &
		\includegraphics[scale=1.5]{rectangulation12} &
		\includegraphics[scale=1.5]{rectangulation13} &
		\\
		$(-2, 5, -9, 6)$ &
		$(4, -9, -1, 6)$ &
		$(-6, 9, 1, -4)$ &
		$(-4, 1, 9, -6)$ &
		$(2, -9, 9, -2)$ &
		$(-2, 9, -9, 2)$ &
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation14} &
		\includegraphics[scale=1.5]{rectangulation15} &
		\includegraphics[scale=1.5]{rectangulation16} &
		\includegraphics[scale=1.5]{rectangulation17} &
		\includegraphics[scale=1.5]{rectangulation19} &
		\includegraphics[scale=1.5]{rectangulation20} &
		\\
		$(6, -9, -1, 4)$ &
		$(4, -1, -9, 6)$ &
		$(-4, 9, 1, -6)$ &
		$(2, -5, 9, -6)$ &
		$(6, -9, 5, -2)$ &
		$(6, -1, -9, 4)$ &
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation21} &
		\includegraphics[scale=1.5]{rectangulation22} &
		\includegraphics[scale=1.5]{rectangulation23} &
		\includegraphics[scale=1.5]{rectangulation24} &
		\includegraphics[scale=1.5]{rectangulation7} &
		\includegraphics[scale=1.5]{rectangulation18}
		\\
		$(-1, 9, -2, -6)$ &
		$(6, -5, 5, -6)$ &
		$(6, 2, -9, 1)$ &
		$(6, 2, -2, -6)$ &
		$(1, -9, 9, -1)$ &
		$(-1, 9, -9, 1)$
	\end{tabular}
	}
	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
	\label{tab:verticesWeakRectangulotope}
\end{table}

>>>>

\begin{theorem}
\label{thm:weakRectangulotope}
  The weak rectangulotope is realized by the polytope~$\WRP (n) \subset \R^n\!$ defined equivalently~as
  \begin{enumerate}[(i)]
  \item the convex hull of the points
  $\sum_{i\in [n]} (\loday{w}^R_i - \antiloday{w}^R_i)\cdot \b{e}_i$ for all weak rectangulations $R$ of size~$n$,
  with
  \[
%  \begin{split}
%    \loday{w}^R_i \eqdef & h(T, i)\cdot v(T,i) \\
%    \antiloday{w}^R_i \eqdef & h(S, i)\cdot v(S,i),
%  \end{split}
    \loday{w}^R_i \eqdef h^T_i\cdot v^T_i
    \qquad\text{and}\qquad
    \antiloday{w}^R_i \eqdef h^S_i\cdot v^S_i,
  \]
  where~$S$ and~$T$ are the source and target trees of the rectangulation~$R$, and $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I \text{ interval of } [n]}{I \not\subseteq S \text{ and } I \not\subseteq [n] \ssm S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$.
  \end{enumerate}
\end{theorem}

For instance, the weak rectangulotope~$\WRP(4)$ is illustrated in \cref{fig:weakRectangulotope}.
The coordinates of the vertex corresponding to each weak rectangulation are gathered in \cref{tab:verticesWeakRectangulotope}.
The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsWeakRectangulotope}.

\begin{table}[p]
	\centerline{
	\begin{tabular}{cccccc}
		\includegraphics[scale=1.5]{rectangulation1} &
		\includegraphics[scale=1.5]{rectangulation2} &
		\includegraphics[scale=1.5]{rectangulation3} &
		\includegraphics[scale=1.5]{rectangulation4} &
		\includegraphics[scale=1.5]{rectangulation5} &
		\includegraphics[scale=1.5]{rectangulation6}
		\\
		$(-3, -1, 1, 3)$ &
		$(-3, -1, 5, -1)$ &
		$(-3, 3, -3, 3)$ &
		$(1, -5, 1, 3)$ &
		$(-3, 0, 5, -2)$ &
		$(-3, 5, -3, 1)$
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation8} &
		\includegraphics[scale=1.5]{rectangulation9} &
		\includegraphics[scale=1.5]{rectangulation10} &
		\includegraphics[scale=1.5]{rectangulation11} &
		\includegraphics[scale=1.5]{rectangulation12} &
		\includegraphics[scale=1.5]{rectangulation13}
		\\
		$(-1, 3, -5, 3)$ &
		$(2, -5, 0, 3)$ &
		$(-3, 5, 0, -2)$ &
		$(-2, 0, 5, -3)$ &
		$(1, -5, 5, -1)$ &
		$(-1, 5, -5, 1)$
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation14} &
		\includegraphics[scale=1.5]{rectangulation15} &
		\includegraphics[scale=1.5]{rectangulation16} &
		\includegraphics[scale=1.5]{rectangulation17} &
		\includegraphics[scale=1.5]{rectangulation19} &
		\includegraphics[scale=1.5]{rectangulation20}
		\\
		$(3, -5, 0, 2)$ &
		$(2, 0, -5, 3)$ &
		$(-2, 5, 0, -3)$ &
		$(1, -3, 5, -3)$ &
		$(3, -5, 3, -1)$ &
		$(3, 0, -5, 2)$
		\\[.2cm]
		&		
		\includegraphics[scale=1.5]{rectangulation21} &
		\includegraphics[scale=1.5]{rectangulation22} &
		\includegraphics[scale=1.5]{rectangulation23} &
		\includegraphics[scale=1.5]{rectangulation24} &
		\\
		&
		$(-1, 5, -1, -3)$ &
		$(3, -3, 3, -3)$ &
		$(3, 1, -5, 1)$ &
		$(3, 1, -1, -3)$ &
	\end{tabular}
	}
	\caption{The vertices of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
	\label{tab:verticesWeakRectangulotope}
\end{table}

\begin{table}
	\centerline{
	\begin{tabular}{c|cccccccccccccc}
		$J$ &
		$1$ &
		$2$ &
		$3$ &
		$4$ &
		$12$ &
		$13$ &
		$14$ &
		$23$ &
		$24$ &
		$34$ &
		$123$ &
		$124$ &
		$134$ &
		$234$
		\\
		\hline
		$f_{\weakeq}(J)$ &
		$3$ &
		$5$ &
		$5$ &
		$3$ &
		$4$ &
		$6$ &
		$5$ &
		$5$ &
		$6$ &
		$4$ &
		$3$ &
		$5$ &
		$5$ &
		$3$
	\end{tabular}
	}
	\caption{The facets of the weak rectangulotope~$\WRP(4)$ of \cref{fig:weakRectangulotope}.}
	\label{tab:facetsWeakRectangulotope}
\end{table}

@@

\begin{theorem}
\label{thm:strongRectangulotope}
  The $(n-1)$-dimensional strong rectangulotope is realized by the polytope~$\SRP (n)$ defined equivalently as
  \begin{enumerate}[(i)]
  \item convex hull of the points
  \[
  \sum_{i,j\in [n], i< j} (\yin{w}^R_{i,j} - \yang{w}^R_{i,j})\cdot (\b{e}_i - \b{e}_j),
  \]
   for all strong rectangulations $R$ of size~$n$, with
  \[
%  \begin{split}
%    \yin{w}_{i,j} \eqdef & h(T, i) \cdot cv (T, S, i, j)\cdot h(S, j)\cdot \llbracket \neg i\backslash j \rrbracket \\% \text { if not } i\backslash j, \text{and } 0\text{ otherwise.} \\ % \neg i|j \text{ and } \neg \frac{\ j\ }i ] \\
%    \yang{w}_{i,j} \eqdef & v(S, i) \cdot ch (S, T, i, j)\cdot v(T, j) \cdot \llbracket \neg j\backslash i \rrbracket,
%  \end{split}
    \yin{w}^R_{i,j} \eqdef h^T_i \cdot cv^{T,S}_{i,j}\cdot h^S_ j\cdot \llbracket \neg \, i \, \backslash \, j \rrbracket
    \qquad\text{and}\qquad
    \yang{w}^R_{i,j} \eqdef v^S_i \cdot ch^{S,T}_{i,j}\cdot v^T_j \cdot \llbracket \neg \, j \, \backslash \, i \rrbracket,
  \]
 where:
  \begin{itemize}
  \item $S$ and $T$ are the source and target trees of the rectangulation~$R$,
  \item $h^T_i$ and $v^T_i$ denote the number of leaves in the horizontal and vertical subtrees of $i$ in~$T$,
  \item $ch^{T,S}_{i,j}$ (resp.~$cv^{T,S}_{i,j}$) denote the number of common leaves of the horizontal (resp.~vertical) subtree of $i$ in $T$ and the horizontal (resp.~vertical) subtree of $j$ in $S$,
  \end{itemize}
  \item the intersection of the hyperplane defined by~$\sum_{i \in [n]} x_i = 0$ with the halfspaces defined by~$\sum_{i \in S} x_i \le \#\set{I,J}{I \not\subseteq S \text{ and } J \not\subseteq [n] \ssm S} + \#\set{I,J}{I \not\subseteq [n] \ssm S \text{ and } J \not\subseteq S}$ for all subsets~$\varnothing \ne S \subsetneq [n]$, where~$I,J$ denote nonempty and consecutive intervals of~$[n]$.
  \end{enumerate}
\end{theorem}

@@

For instance, the strong rectangulotope~$\SRP(4)$ is illustrated in \cref{fig:strongRectangulotope}.
The coordinates of the vertex corresponding to each strong rectangulation are gathered in \cref{tab:verticesStrongRectangulotope}.
The facet inequalities corresponding to each nonempty proper subset of~$[n]$ are gathered in \cref{tab:facetsStrongRectangulotope}.

\begin{table}
	\centerline{
	\begin{tabular}{cccccc}
		\includegraphics[scale=1.5]{rectangulation1} &
		\includegraphics[scale=1.5]{rectangulation2} &
		\includegraphics[scale=1.5]{rectangulation3} &
		\includegraphics[scale=1.5]{rectangulation4} &
		\includegraphics[scale=1.5]{rectangulation5} &
		\includegraphics[scale=1.5]{rectangulation6}
		\\
		$(-6, -2, 2, 6)$ &
		$(-6, -2, 9, -1)$ &
		$(-6, 5, -5, 6)$ &
		$(1, -9, 2, 6)$ &
		$(-6, 1, 9, -4)$ &
		$(-6, 9, -5, 2)$
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation8} &
		\includegraphics[scale=1.5]{rectangulation9} &
		\includegraphics[scale=1.5]{rectangulation10} &
		\includegraphics[scale=1.5]{rectangulation11} &
		\includegraphics[scale=1.5]{rectangulation12} &
		\includegraphics[scale=1.5]{rectangulation13}
		\\
		$(-2, 5, -9, 6)$ &
		$(4, -9, -1, 6)$ &
		$(-6, 9, 1, -4)$ &
		$(-4, 1, 9, -6)$ &
		$(2, -9, 9, -2)$ &
		$(-2, 9, -9, 2)$
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation14} &
		\includegraphics[scale=1.5]{rectangulation15} &
		\includegraphics[scale=1.5]{rectangulation16} &
		\includegraphics[scale=1.5]{rectangulation17} &
		\includegraphics[scale=1.5]{rectangulation19} &
		\includegraphics[scale=1.5]{rectangulation20}
		\\
		$(6, -9, -1, 4)$ &
		$(4, -1, -9, 6)$ &
		$(-4, 9, 1, -6)$ &
		$(2, -5, 9, -6)$ &
		$(6, -9, 5, -2)$ &
		$(6, -1, -9, 4)$
		\\[.2cm]
		\includegraphics[scale=1.5]{rectangulation21} &
		\includegraphics[scale=1.5]{rectangulation22} &
		\includegraphics[scale=1.5]{rectangulation23} &
		\includegraphics[scale=1.5]{rectangulation24} &
		\includegraphics[scale=1.5]{rectangulation7} &
		\includegraphics[scale=1.5]{rectangulation18}
		\\
		$(-1, 9, -2, -6)$ &
		$(6, -5, 5, -6)$ &
		$(6, 2, -9, 1)$ &
		$(6, 2, -2, -6)$ &
		$(1, -9, 9, -1)$ &
		$(-1, 9, -9, 1)$
	\end{tabular}
	}
	\caption{The vertices of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
	\label{tab:verticesStrongRectangulotope}
\end{table}

\begin{table}
	\centerline{
	\begin{tabular}{cccccccc}
	\end{tabular}
	}
	\caption{The facets of the strong rectangulotope~$\SRP(4)$ of \cref{fig:strongRectangulotope}.}
	\label{tab:facetsStrongRectangulotope}
\end{table}
